---
title: "STA463Project2"
author: "Steve, Jairus, Allen"
date: "2023-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(GGally)
library(ggfortify)
library(car)
```


## I. Introduction  

 This dataset shows Airbnb prices in Amsterdam during weekdays. Each listing is evaluated for room types, cleanliness, and distance from the city center.  
The dataset includes :
Response - price: The total price of the listing 
Predictor 1 - room_type: The type of room being offered 
Predictor 2 - dist: Distance from the city center 
Predictor 3 - cleanliness_rating: The cleanliness rating of the listing 
Link : https://www.kaggle.com/datasets/thedevastator/airbnb-prices-in-european-cities?select=amsterdam_weekdays.csv

<<<<<<< HEAD
yes
  
=======
>>>>>>> 8592249a7217e60a52ba26e9f31560ea2353ca11
## II. Exploratory Data Analysis  
 
```{r, echo=FALSE}
plc <- read.csv("placement.csv")
summary(plc)
noquote(paste('Correlation:',round(cor(plc$cgpa, plc$placement_exam_marks), 2)))
```

```{r, echo=FALSE, message=FALSE}
ggplot(data=plc, aes(x=cgpa, y=placement_exam_marks)) +
  geom_point() +
  xlab("CGPA") +
  ylab("Placement Exam Marks") +
  geom_smooth(method="lm", se=FALSE)
```
  
## III. Model and Analysis

## IV. Conclusion
