---
title: "STA463Project2"
author: "Steve, Jairus, Allen"
date: "2023-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(GGally)
library(ggfortify)
library(car)
```


## I. Introduction  

This project explores a dataset pertaining to Airbnb prices in Amsterdam during weekdays. Each Airbnb listing is an observation, and is evaluated for room types, cleanliness, and distance from Amsterdam's city center.

Our variables under consideration:

* Y = price (Euros) 
* X1 = room_type (categorical: Private room, Entire home/apt)
* X2 = dist (distance from city center in kilometers)
* X3 = cleanliness_rating (numerical, ranging from 4 to 10)

We will first identify the structure of our dataset:

```{r, echo=FALSE}
airbnb <- read.csv("airbnb.csv")
head(airbnb)
```

This project involves performing a confirmatory multiple regression analysis. Specifically, we seek to answer the following research question: Does the relationship between Airbnb prices in Amsterdam and distance from the city center differ for Airbnbs in private rooms versus Airbnbs in the entire home or apartment?

=======

## II. Exploratory Data Analysis  
 



## IV. Conclusion
